---
title: "Video Game Sales"
author: "Jing Li"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    toc: yes
  html_notebook:
    theme: lumen
    toc: yes
---

# Import Data

The dataset was downloaded from [kaggle.com](https://www.kaggle.com/gregorut/videogamesales). 

```{r, setup, echo=FALSE, message=FALSE}
options(width = 119)
library(readr)
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(ggplot2)
library(ggrepel)
library(pander)
options("readr.num_columns" = 0)
vgsales <- read_csv("vgsales.csv")
```

# Data Overview

There are __11__ variables and __16,598__ observations.

The variables are:

* `Rank`: Ranking of overall sales
* `Name`: The game name
* `Platform`: Platform of the games release (i.e. PC,PS4, etc.)
* `Year`: Year of the game's release
* `Genre`: Genre of the game
* `Publisher`: Publisher of the game
* `NA_Sales`: Sales in North America (in millions)
* `EU_Sales`: Sales in Europe (in millions)
* `JP_Sales`: Sales in Japan (in millions)
* `Other_Sales`: Sales in the rest of the world (in millions)
* `Global_Sales`: Total worldwide sales (in millions)

_(sales are the number of copies)_

Below are the summary statistics for the numerical variables:

```{r, message=FALSE}
vgsales %>% 
  select(dplyr::contains("_Sales")) %>% 
  summary() %>% 
  pandoc.table(style = "simple")
```

# Data Visualizations

## 1. Genre Popularity over the Years
```{r, fig.width=9}
vgsales %>% 
  ggplot(aes(x = Year, y = Global_Sales, fill = Genre)) +
  geom_bar(stat = "identity") +
  labs(title = "Genre Popularity over Years", x = "Year", y = "Global Sales (in milion copies)") +
  scale_x_discrete(breaks = as.character(seq(from = 1980, to = 2016, by = 4))) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_brewer(palette = "Set3")
```

It seems that __Action__ game is the dominating genre since 2000, and __Sports__ game is usually the second popular genre.
Also, the global sales peaked around __2008__ and then started to drop dramatically. 

## 2. Genre Sales by Regions

```{r, fig.width=9}
vgsales %>% 
  group_by(Genre) %>% 
  summarise_if(is.numeric, sum) %>% 
  ungroup() %>% 
  gather(Region, Sales, NA_Sales:Global_Sales) %>% 
  mutate(Region = str_replace(Region, "_Sales$", "") %>% 
           factor(levels = c("NA", "EU", "JP", "Other", "Global"),
                  labels = c("North America", "Europe", "Japan", "Other", "Global"))) %>% 
  ggplot(aes(x = Region, y = Genre)) +
  geom_tile(aes(fill = Sales)) +
  labs(title = "Genre Sales by Regions", x = "Region", y = "Genre") +
  scale_fill_continuous(name = "Sales\n(in million copies)") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), legend.title.align = 0)
```

The sales numbers are aggregated to remove the time dimension. From the graph above, we can see that in terms of sales, people from `North America` and those from `Europe` have very similar tastes as they all prefer __Action__, __Sports__, __Shooter__, and 
__Platform__ over others, though in terms of absolute numbers, games are sold more in `North America` than in `Europe`. 

However, in `Japan`, the most popular genre is __Role-Playing__, which might be due to the cultural differences.

Globally, __Action__ and __Sports__ are the most popular genres.

## 3. Global Sales by Genre of the Top 10 Publishers

Below shows the __Top 10__ Publishers in terms of cumulative global sales over the years:
```{r}
top10 <- vgsales %>% 
  group_by(Publisher) %>% 
  summarise(Cum_Sales = sum(Global_Sales)) %>% 
  arrange(desc(Cum_Sales)) %>% 
  top_n(10)
top10 %>% pandoc.table(style = "rmarkdown")
```

```{r, fig.width=9}
publishers <- top10$Publisher
vgsales %>% 
  filter(Publisher %in% publishers) %>% 
  ggplot(aes(x = Genre, y = Global_Sales)) +
  geom_bar(stat = "identity", aes(fill = Publisher)) +
  coord_flip() +
  labs(title = "Global Sales by Top 10 Publisher by Genre", xlab = "Cumulative Global Sales") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_brewer(palette = "Set3")
```

Within the Top 10 publishers, __Sports__ games have the greatest global sales, followed by __Action__ games. 
__Electronic Arts__ has absolute power over the __Sports__ games, while __Take-Two Interactive__ is the No. 1 __Action__
game publisher. Although __Puzzle__, __Role Playing__ and __Platform__ games don't have the most sales, in those worlds, 
__Nintendo__ reigns. And __Shooter__ game is __Activision__'s playground.

## 4. Top 5 Publishers (Global Sales) over the Decades

The graph below shows how the __Top 5 Publishers__ change over time. However, it's worth noting that __2010s__ doesn't have complete data as 
we are still in the middle of it!
```{r, fig.width=10}
top5bydecades <- vgsales %>% 
  mutate(decade = cut(as.numeric(Year), breaks = c(1980, 1990, 2000, 2010, 2020), 
                      labels = c("1980s", "1990s", "2000s", "2010s"), ordered_result = TRUE)) %>% 
  filter(!is.na(decade)) %>% 
  group_by(decade, Publisher) %>% 
  summarise(Cum_Sales = sum(Global_Sales)) %>% 
  arrange(decade, desc(Cum_Sales)) %>% 
  top_n(5)

top5bydecades %>% 
  ggplot(aes(x = decade, y = Cum_Sales)) +
  geom_label_repel(aes(label = Publisher), label.size = 0.2) +
  labs(title = "Top 5 Publishers over the Decades", x = "Decade", y = "Cumulative Sales") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
  
```

Legendarily, __Nintendo__ is the only one that has been in this short list for 4 decades. Founded in 1982, __Electronic Arts__ since then has been climbing on this list step by step, and 
finally in the current decade becomes the __No.1__ game publisher as of 2016.

Also, the gaming industry seems to have reached the apex during the __2000s__, even though we are only half way through the current decade, the absolute difference between __2000s__ and __2010s__ is still significantly large.

## 5. XBox vs PlayStation

```{r, fig.width=10, fig.height=10}
xbox_ps <- vgsales %>% 
  filter(Platform %in% c("XB", "X360", "XOne", "PS", str_c("PS", 2:4))) %>% 
  mutate(Platform = plyr::mapvalues(Platform, from = c("XB", "X360", "XOne", "PS", str_c("PS", 2:4)),
                                    to = c(str_c("XBOX", c("", " 360", " One")), str_c("PlayStation ", 1:4)))) %>% 
  group_by(Year, Platform) %>% 
  summarise_if(is.numeric, sum) %>% 
  ungroup() %>% 
  gather(Region, Sales, NA_Sales:Global_Sales) %>% 
  mutate(Region = str_replace(Region, "_Sales$", "") %>% 
           factor(levels = c("NA", "EU", "JP", "Other", "Global"),
                  labels = c("North America", "Europe", "Japan", "Other", "Global"))) %>% 
  filter(!str_detect(Year, "N/A|2017"))

xbox_ps %>% 
  ggplot(aes(x = Year, y = Sales)) +
  geom_bar(aes(fill = Platform), stat = "identity") +
  facet_grid(Region~., scales = "free") +
  theme_minimal() +
  scale_fill_manual(values = c("PlayStation 1" = "#F7E8C3", "PlayStation 2" = "#EF7079", "PlayStation 3" = "#BF382A", "PlayStation 4" = "#683531", 
                               "XBOX" = "#D5EEFF", "XBOX 360" = "#007CB9", "XBOX One" = "#005689"))
```

As two very (if not the most) famous gaming consols, __XBOX__ and __PlayStation__ have been competing against each other for over a decade.

From the graph above, most of the __PlayStation__ game sales come from Japan, the home of the brand, while since its release in 2001, the popularity of 
__XBOX__ has increased a lot global-wise. In __North America__, __XBOX__ and __PlayStation__ seem to be taking up
the market share equally, but everywhere else, __PlayStation__ has relatively more share and in Japan, __PlayStation__ is just pure monopoly.

Though introduced at around the same time, __PlayStation__ tends to outperform __XBOX One__ in every market, in terms of game sales. 
