---
title: "Video Game Sales"
author: "Jing Li"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_notebook:
    toc: true
    theme: lumen
---

# import data

The dataset was downloaded from [kaggle.com](https://www.kaggle.com/gregorut/videogamesales). 

```{r, echo=FALSE, message=FALSE}
options(width = 119)
library(readr)
options("readr.num_columns" = 0)
vgsales <- read_csv("vgsales.csv")
```

# data overview

There are __11__ variables and __16,598__ observations.

The variables are:

* `Rank`: Ranking of overall sales
* `Name`: The game name
* `Platform`: Platform of the games release (i.e. PC,PS4, etc.)
* `Year`: Year of the game's release
* `Genre`: Genre of the game
* `Publisher`: Publisher of the game
* `NA_Sales`: Sales in North America (in millions)
* `EU_Sales`: Sales in Europe (in millions)
* `JP_Sales`: Sales in Japan (in millions)
* `Other_Sales`: Sales in the rest of the world (in millions)
* `Global_Sales`: Total worldwide sales (in millions)

_(sales are the number of copies)_

Below are the summary statistics for the numerical variables:

```{r, message=FALSE}
library(dplyr)
vgsales %>% 
  select(contains("_sales")) %>% 
  summary()
```

